---
title: "paper"
output: html_document
date: "2025-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 200)
require(tidyverse)
require(papaja)
require(RColorBrewer)
require(tools)
require(cowplot)
```

```{r data, include=FALSE}
words = read_csv('data/kidwords/kidwords.csv', col_names = c('word'))

# model summaries
model_summaries = read_csv('outputs/brute_force_2_with_frequency/model_summaries.csv') %>% 
  pivot_longer(cols = accuracy_train:mse_holdout, names_to = "variable", values_to = "value") %>% 
  mutate(condition = case_when(str_detect(variable, "train") ~ "train",
                               str_detect(variable, "test") ~ "test",
                               str_detect(variable, "holdout") ~ "holdout"),
         metric = str_split(variable, "_", simplify = T)[,1]) %>% 
  filter(condition != "test") %>% 
  filter((hidden_units %in% c(20, 30, 40))) %>% 
  select(-variable)


model_summaries = model_summaries %>% 
  left_join(model_summaries %>% 
              group_by(metric) %>% 
              summarise(M = mean(value),
                        SD = sd(value))) %>% 
  mutate(value_z = (value-M)/SD)

library(broom)


#model_summaries %>% 
#  write_csv("~/Desktop/the_words_matter_model_summaries.csv")

```

# Methods
## Frequencies
```{r}

probabilities = read_csv('word_frequencies_and_probabilities.csv') %>% 
  glimpse()

probabilities %>% 
  arrange(desc(frequency)) %>% 
  head()


probabilities %>% 
  arrange(frequency) %>% 
  view()

```


# Results
## Differences in outcomes on training and generalization words
Means, standard deviations, and extreme values for raw and standardized MSE and raw accuracy for the train and generalization sets. Here we analyze data without respect to (collapsing across) representational capacity.
```{r chunk1}
model_summaries %>% 
  filter(metric == "mse" | metric == "accuracy") %>% 
  group_by(condition, metric) %>% 
  filter(metric == "mse") %>% 
  summarise(mse_mean = mean(value),
            mse_sd = sd(value),
            mse_z_mean = mean(value_z),
            mse_z_sd = sd(value_z),
            mse_min = min(value),
            mse_z_min = min(value_z),
            mse_max = max(value),
            mse_z_max = max(value_z))

```

### Top and bottom performers

```{r}

bottom_100 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(condition) %>%
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  pull(model_id)

model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  filter(model_id %in% bottom_100)

# correlation
tmp  = model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank_train = seq_len(n())) %>% 
  select(tmp_id, rank_train)


tmp_2 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank_holdout = seq_len(n())) %>% 
  left_join(tmp) 

cor.test(tmp_2$rank_holdout, tmp_2$rank_train)

```

### Figure 1

```{r chunk3}
plot = model_summaries %>% 
  filter(metric == "mse") %>% 
  mutate(condition = factor(condition, levels = c("train", "holdout"))) %>% 
  ggplot(aes(factor(condition), value_z, fill = factor(condition))) +
  geom_violin() +
  geom_boxplot(width = .2) +
  theme_apa() +
  labs(x = "Set", y = "Mean squared error (Z)") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("train" = "Train", "holdout" = "Generalization")) +
  scale_fill_manual(values = c("goldenrod3", "firebrick4"))


ggsave("results_figure1.png", plot = plot, width = 5, height = 4, dpi = 600)

```


### Table 3

```{r}
model_summaries %>% 
  filter(metric == "mse") %>% 
  group_by(condition, hidden_units) %>% 
  summarise(mse_mean = mean(value),
          mse_sd = sd(value),
          mse_z_mean = mean(value_z),
          mse_z_sd = sd(value_z),
          mse_min = min(value),
          mse_z_min = min(value_z),
          mse_max = max(value),
          mse_z_max = max(value_z)) %>% 
  mutate(MSE = str_c(mse_mean, " (", mse_z_mean, ")"),
         SD = str_c(mse_sd, " (", mse_z_sd, ")"),
         MIN = str_c(mse_min, " (", mse_z_min, ")"),
         MAX = str_c(mse_max, " (", mse_z_max, ")")) %>% 
  select(hidden_units, MSE, SD, MIN, MAX) %>% 
  write_csv("~/Desktop/table3.csv")


```

### Figure 2

```{r}


plot = model_summaries %>% 
  filter(metric == "mse") %>% 
  mutate(condition = case_when(condition == "holdout" ~ "Generalization",
                               condition == "train" ~ "Train")) %>% 
  mutate(condition = factor(condition, levels = c("Train", "Generalization"))) %>% 
  ggplot(aes(factor(hidden_units), value_z, fill = interaction(factor(condition), factor(hidden_units)))) +
  geom_violin() +
  geom_boxplot(width = .5) +
  theme_apa() +
  facet_wrap(~condition) +
  theme(legend.position = "none") +
  labs(x = "Representational capacity (hidden units)",
       y = "Mean squared error (Z)") +
  scale_fill_manual(values = c("Train.20" = "lightgoldenrod", "Train.30" = "goldenrod3", "Train.40" = "orange", "Generalization.20" = "lightcoral", "Generalization.30" = "firebrick4", "Generalization.40" = "deeppink4"))

ggsave("results_figure2.png", plot = plot, width = 5, height = 4, dpi = 600)


```

### Differences in Specific Ensembles at Different Levels of Representational Capacity
The correlation between 20 and 40 hidden units.

```{r}
tmp = model_summaries %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 20 & condition == "holdout") %>% 
  select(model_id, hu20 = value_z) %>% 
  left_join(model_summaries %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 40 & condition == "holdout") %>% 
  select(model_id, hu40 = value_z))

cor.test(tmp$hu20, tmp$hu40)

```

The correlation across all levels of hidden unit (10 to 100 in increments of 10)
```{r}

model_summaries_10_to_100 = read_csv('outputs/brute_force_2_with_frequency/model_summaries.csv') %>% 
  pivot_longer(cols = accuracy_train:mse_holdout, names_to = "variable", values_to = "value") %>% 
  mutate(condition = case_when(str_detect(variable, "train") ~ "train",
                               str_detect(variable, "test") ~ "test",
                               str_detect(variable, "holdout") ~ "holdout"),
         metric = str_split(variable, "_", simplify = T)[,1]) %>% 
  filter(condition != "test") %>% 
  filter((hidden_units %in% c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100))) %>% 
  select(-variable) 


model_summaries_10_to_100 = model_summaries_10_to_100 %>% 
  left_join(model_summaries_10_to_100 %>% 
              group_by(metric) %>% 
              summarise(M = mean(value),
                        SD = sd(value))) %>% 
  mutate(value_z = (value-M)/SD)



tmp = model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 10 & condition == "holdout") %>% 
  select(model_id, hu10 = value_z) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 20 & condition == "holdout") %>% 
  select(model_id, hu20 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 30 & condition == "holdout") %>% 
  select(model_id, hu30 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 40 & condition == "holdout") %>% 
  select(model_id, hu40 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 50 & condition == "holdout") %>% 
  select(model_id, hu50 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 60 & condition == "holdout") %>% 
  select(model_id, hu60 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 70 & condition == "holdout") %>% 
  select(model_id, hu70 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 80 & condition == "holdout") %>% 
  select(model_id, hu80 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 90 & condition == "holdout") %>% 
  select(model_id, hu90 = value_z)) %>% 
  left_join(model_summaries_10_to_100 %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 100 & condition == "holdout") %>% 
  select(model_id, hu100 = value_z))


require(psych)
pairs.panels(tmp[, 2:ncol(tmp)]) 

cor(tmp[, 2:ncol(tmp)])

tmp %>% 
  left_join(read_csv('../words/elp/elp_full_5.27.16.csv') %>% 
  select(word = Word, naming_rt = I_NMG_Mean_RT, naming_accuracy = I_NMG_Mean_Accuracy))
```


#### UFO Plot

```{r}
text_size = 6

differences_in_rank = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(hidden_units) %>%
  #summarise(acc = mean(value_z)) %>% 
  arrange(desc(value_z)) %>% 
  mutate(rank = seq_len(n())) %>% 
  ungroup() %>%
  group_by(model_id) %>%
  arrange(-desc(hidden_units)) %>% 
  summarize(difference = diff(rank))


plot_a = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  left_join(differences_in_rank, by = "model_id") %>% 
  ggplot(aes(factor(hidden_units), value_z)) +
  geom_line(size = .2, aes(group = model_id, color = difference)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .2, outlier.size = .2) +
  theme_apa() +
  labs(title = "Differences in rank outcome",
       subtitle = "20, 40 hidden units (holdout only)",
       x = "Representational capacity (in hidden units)", 
       y = "Mean squared error (Z)", 
       fill = "Hidden units") +
  scale_fill_manual(values = c("lightcoral", "deeppink4")) +
  scale_color_gradientn(colors = heat.colors(10)) +
  theme(legend.position = "none",
        plot.title = element_text(size = text_size),
        plot.subtitle = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size))

plot_a


plot_b = differences_in_rank %>% 
  ggplot(aes(difference)) +
  geom_histogram(color = "black", fill = "turquoise") +
  labs(x = "Difference in rank outcome", y = "Count",
       title = "Differences in rank outcome for each model") +
  theme_apa() +
  theme(plot.title = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size))
  
plot_b  

# spotlighting top movers
bottom_10 = differences_in_rank %>% 
  arrange(difference) %>% 
  slice_head(n = 10) %>% 
  mutate(status = "bottom")

top_bottom_10 = differences_in_rank %>% 
  arrange(desc(difference)) %>% 
  slice_head(n = 10) %>% 
  mutate(status = "top") %>% 
  rbind(bottom_10)

plot_c = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  left_join(top_bottom_10)  %>%
  mutate(model_id_ = case_when(is.na(status) ~ NA,
                               !is.na(status) ~ model_id)) %>% 
  ggplot(aes(factor(hidden_units), value_z)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .2, outlier.size = .2) +
  geom_line(size = .2, na.rm = TRUE, alpha = .9, aes(group = model_id_, color = status, linetype = status)) +
  theme_apa() +
  labs(title = "Top 20 movers in differences in outcome",
       subtitle = "models at 20, 40 hidden units (holdout only)",
       x = "Representational capacity (in hidden units)", 
       y = "Mean squared error (Z)", 
       fill = "Hidden units") +
  scale_fill_manual(values = c("lightcoral", "deeppink4")) +
  #scale_color_gradientn(colors = heat.colors(10)) +
  theme(legend.position = "none",
        plot.title = element_text(size = text_size),
        plot.subtitle = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size)) +
  scale_color_manual(values = c("top" = "grey44", "bottom" = "black")) +
  scale_linetype_manual(values = c("dashed", "dotted"))


plot_c

top<- plot_grid(plot_a, plot_c, ncol = 2)

# Arrange the bottom row plot centered
bottom <- plot_grid(NULL, plot_b, NULL, ncol = 3, rel_widths = c(.5, 1.5, .5))

# Combine the top and bottom rows
plot <- plot_grid(top, bottom, ncol = 1, rel_heights = c(1, 1))
plot


ggsave("results_figure3.png", plot = plot, width = 5, height = 4, dpi = 600)
```


