---
title: "TDF2024"
output: html_document
date: "2024-04-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(papaja)
require(RColorBrewer)
require(tools)
```

```{r read, include=FALSE}
top_bottom_20_ids = read_csv('data/top_and_bottom_20_learners_10_and_100_hidden_units.csv') %>% 
  distinct(model_id) %>% 
  pull(model_id)

holdouts = read_csv('data/holdouts.csv', col_names = F)[,1]

holdouts = holdouts %>% 
  mutate(word = words$word) %>% 
  rename(holdout = X1)

words = read_csv('data/kidwords/kidwords.csv', col_names = c('word')) %>% 
  mutate(holdout = holdouts$holdout)

train = read_csv('data/samples.csv', col_names = str_c("set_", seq(0, 9999)))

top_bottom_20_train = train %>% 
  select(all_of(str_c("set_", top_bottom_20_ids)))

filenames <- list.files(path = "outputs/top_and_bottom_20_learners/", pattern = "accuracies\\.csv$", full.names = TRUE)


top_bottom_20_test_list = list()

for (i in seq(1, ncol(top_bottom_20_train))){
  
  colname = names(top_bottom_20_train[,i])
  sample = str_extract(colname, "(?<=set_).*")
  
  df = top_bottom_20_train[,i] %>% 
    mutate(word = words$word,
           holdout = holdouts$holdout,
           model_id = as.numeric(sample)) %>% 
    rename(train = all_of(colname))
  
  top_bottom_20_test_list[[as.character(i)]] = df
  
}

top_bottom_20_train_holdout = list_rbind(top_bottom_20_test_list)

top_bottom_20_list = list()

for (filename in filenames){
    
    model_id = str_extract(filename, paste("(?<=sample_).*(?=_hidden_)"))
    hidden_units = str_extract(filename, "(?<=hidden_).*?(?=_)")
    epoch = str_extract(filename, "(?<=_)[0-9]+(?=__)")

    df = read_csv(filename, col_names = c("accuracy")) %>% 
        mutate(word = words$word,
               model_id = as.numeric(model_id),
               hidden_units = as.numeric(hidden_units),
               epoch = as.numeric(epoch))
      
      top_bottom_20_list[[filename]] = df
    } 


top_bottom_20 = list_rbind(top_bottom_20_list) %>% 
  left_join(read_csv('data/top_and_bottom_20_learners_10_and_100_hidden_units.csv'), by = c("model_id", "hidden_units")) %>% 
  left_join(top_bottom_20_train_holdout, by = c("model_id", "word"))



# model summaries
TDF2024 = read_csv('outputs/brute_force_2_with_frequency/model_summaries.csv') %>% 
  pivot_longer(cols = accuracy_train:mse_holdout, names_to = "variable", values_to = "value") %>% 
  mutate(condition = case_when(str_detect(variable, "train") ~ "train",
                               str_detect(variable, "test") ~ "test",
                               str_detect(variable, "holdout") ~ "holdout"),
         metric = str_split(variable, "_", simplify = T)[,1]) %>% 
  select(-variable)


TDF2024 = TDF2024 %>% 
  left_join(TDF2024 %>% 
              group_by(metric) %>% 
              summarise(M = mean(value),
                        SD = sd(value))) %>% 
  mutate(value_z = (value-M)/SD)
  


msd = TDF2024 %>% 
    filter(metric == "accuracy") %>% 
    summarise(M = mean(value),
            SD = sd(value))

top_bottom_20 = top_bottom_20 %>% 
  mutate(M = msd$M,
         SD = msd$SD,
         accuracy_z = (accuracy-M)/SD)



```


## For poster
# Plot 1
Differences between 20 hidden unit and 100 hidden unit conditions in terms of standardized accuracy
```{r}
plot_1 = TDF2024 %>% 
  filter(hidden_units %in% c(10, 40, 70, 100)) %>% 
  filter(metric == "accuracy") %>% 
  mutate(Condition = toTitleCase(condition)) %>% 
  filter(Condition %in% c("Holdout", "Train")) %>% 
  mutate(Condition = as.factor(Condition)) %>% 
  mutate(Condition = fct_relevel(Condition, "Train", "Holdout")) %>% 
  ggplot(aes(factor(hidden_units), value_z)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .3, outlier.shape = NA) +
  facet_wrap(vars(Condition)) +
  theme_apa() +
  labs(title = "Differences in learning of ensembles of words across learners with different skill",
       x = "Representational capacity (in hidden units)", 
       y = "Accuracy standardized (one point per model)", fill = "Hidden units") +
  scale_fill_brewer(palette = "Spectral") +
  theme(legend.position = "none",
        axis.title = element_text(size = 28),
        axis.text = element_text(size = 22),
        strip.text = element_text(size = 28),
        plot.title = element_text(size = 30))

plot_1
ggsave("plot_1.png", plot = plot_1)

```

## Plot 2
Magnitude of differences between best and worst performers across hidden units.

```{r}
# compute the differences so we can use them for plot 2a
plot_2b_data =  TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>%
  group_by(hidden_units, model_id) %>%
  summarise(acc = mean(value_z)) %>% 
  arrange(desc(acc)) %>% 
  mutate(acc_rank = seq_len(n())) %>% 
  ungroup() %>%
  group_by(model_id) %>%
  arrange(-desc(hidden_units)) %>% 
  summarize(difference = diff(acc_rank))


plot_2a_alt = TDF2024 %>% 
  filter(hidden_units %in% c(10, 40, 70, 100)) %>% 
  filter(metric == "accuracy") %>% 
  mutate(Condition = toTitleCase(condition)) %>% 
  filter(Condition %in% c("Train")) %>% 
  group_by(hidden_units) %>% 
  summarise(acc_min = min(value_z),
            acc_max = max(value_z)) %>%
  ungroup() %>% 
  mutate(acc_diff = acc_max - acc_min) %>% 
  ggplot(aes(factor(hidden_units), acc_diff, fill = factor(hidden_units))) +
  geom_bar(stat = "identity", color = "black") +
  theme_apa() +
  labs(title = "Differences in learning extremes at each level of hidden units",
       x = "Representational capacity (in hidden units)", 
       y = "Difference between highest + lowest learner (in SD units) ", 
       fill = "Hidden units") +
  scale_fill_brewer(palette = "Spectral") +
  theme(legend.position = "none",
        axis.title = element_text(size = 28),
        axis.text = element_text(size = 22),
        strip.text = element_text(size = 28),
        plot.title = element_text(size = 30))

plot_2a = TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>% 
  mutate(Condition = toTitleCase(condition)) %>% 
  group_by(hidden_units, model_id) %>% 
  summarise(acc = mean(value_z)) %>%
  left_join(plot_2b_data, by = "model_id") %>% 
  ggplot(aes(factor(hidden_units), acc)) + #, fill = factor(hidden_units)
  geom_line(size = .05, aes(group = model_id, color = difference)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .2) +
  theme_apa() +
  labs(title = "Differences in learning at 10, 100 hidden units",
       x = "Representational capacity (in hidden units)", 
       y = "Accuracy (in SDs - one point per model)", 
       fill = "Hidden units") +
  scale_fill_manual(values = c("firebrick", "#0072B2")) +
  scale_color_gradientn(colors = heat.colors(10)) +
  theme(legend.position = "none",
        axis.title.x = element_text(size = 28),
        axis.title.y = element_text(size = 24),
        axis.text = element_text(size = 22),
        strip.text = element_text(size = 28),
        plot.title = element_text(size = 32))

plot_2a
ggsave("plot_2a.png", plot = plot_2a)





plot_2b = plot_2b_data %>% 
  ggplot(aes(difference)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey60") +
  geom_density(size = 1, color = "brown4") +
  geom_vline(xintercept = mean(plot_2b_data$difference), linetype = "dashed") +
  theme_apa() +
  theme(legend.position = "none") +
  labs(title = "Differences in rank accuracy per model", subtitle = "(10 v. 100 hidden unit models)",
       x = "Difference in rank",
    y = "Density") +
  theme(legend.position = "none",
        axis.title = element_text(size = 28),
        axis.text = element_text(size = 22),
        strip.text = element_text(size = 28),
        plot.title = element_text(size = 30),
        plot.subtitle = element_text(size = 28))

plot_2b
ggsave("plot_2b.png", plot = plot_2b)
```


## Correlations between 10 and 100, extension of plot 2

```{r}

TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>% 
  #filter(condition != "test") %>% 
  group_by(hidden_units, model_id) %>% 
  summarise(acc = mean(value_z)) %>% 
  pivot_wider(names_from = hidden_units, values_from = acc) %>% 
  rename(hu10 = `10`, hu100 = `100`) %>% 
  summarise(cor(hu10, hu100))


TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>% 
  #filter(condition != "test") %>% 
  group_by(hidden_units, model_id) %>% 
  summarise(acc = mean(value_z)) %>% 
  pivot_wider(names_from = hidden_units, values_from = acc) %>% 
  rename(hu10 = `10`, hu100 = `100`) %>% 
  ggplot(aes(hu10, hu100)) +
  geom_point() +
  geom_smooth() +
  theme_apa()


```


## Plot 3a + 3b
Variation of words in holdout.

```{r}
COLORS = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot_3a = top_bottom_20 %>% 
  filter(holdout == 1) %>% 
  filter(epoch == 49) %>% 
  group_by(hidden_units, word) %>% 
  summarise(acc = mean(accuracy_z)) %>% 
  ggplot(aes(factor(hidden_units), acc, color = factor(hidden_units), group = word)) +
  geom_line(color = "grey44", alpha = .4) +
  geom_point(position = position_jitter(width = .03)) +
  theme_apa() +
    theme(legend.position = "none") +
  labs(title = "Accuracy across two levels of hidden units",
       x = "Representational capacity (low - high)",
    y = "Accuracy standardized (per word)") +
  scale_color_manual(values = c("firebrick", "#0072B2")) +
  theme(legend.position = "none",
        axis.title = element_text(size = 28),
        axis.text = element_text(size = 22),
        plot.title = element_text(size = 30))

plot_3a
ggsave(filename = "plot_3a.png", plot = plot_3a)

plot_3b_data = top_bottom_20 %>% 
  filter(holdout == 1) %>% 
  filter(epoch == 49) %>% 
  group_by(hidden_units, word) %>% 
  summarise(acc = mean(accuracy_z)) %>% 
  arrange(desc(acc)) %>% 
  mutate(acc_rank = seq_len(n())) %>% 
  ungroup() %>% 
  group_by(word) %>%
  arrange(-desc(hidden_units)) %>% 
  summarize(difference = diff(acc_rank))

plot_3b = plot_3b_data %>% 
  ggplot(aes(difference)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey44") +
  geom_density(color = "brown4", size = 1) + 
  geom_vline(xintercept = mean(plot_3b_data$difference), linetype = "dashed") +
  theme_apa() +
  theme(legend.position = "none") +
  labs(title = "Differences in rank accuracy per word", subtitle = "(10 v. 100 hidden unit models)",
       x = "Difference in rank",
    y = "Density") +
  theme(legend.position = "none",
        axis.title = element_text(size = 28),
        axis.text = element_text(size = 22),
        plot.title = element_text(size = 30),
        plot.subtitle = element_text(size = 28))


plot_3b
ggsave(filename = "plot_3b.png", plot = plot_3b)

```


## Highest and lowest performers at 10 and 100 hidden units

```{r top_and_bottom_learners_10_100}

TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>% 
  group_by(hidden_units) %>% 
  arrange(desc(value)) %>% 
  reframe(acc_rank = seq_len(n()),
          model_id = model_id) %>% 
  filter(acc_rank %in% c(1:20) | acc_rank %in% c(9981:10000)) %>% 
  mutate(condition = case_when(acc_rank %in% c(1:20) ~ "top",
                               acc_rank %in% c(9981:10000) ~ "bottom",
                               TRUE ~ NA)) %>% 
  select(model_id, hidden_units, condition) %>% 
  write_csv('data/top_and_bottom_20_learners_10_and_100_hidden_units.csv')
```


```{r}
TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  filter(metric == "accuracy") %>% 
  group_by(hidden_units) %>% 
  arrange(desc(value)) %>% 
  reframe(acc_rank = seq_len(n()),
          model_id = model_id,
          value_z) %>% 
  filter(acc_rank %in% c(1:20) | acc_rank %in% c(9981:10000)) %>% 
  mutate(condition = case_when(acc_rank %in% c(1:20) ~ "top",
                               acc_rank %in% c(9981:10000) ~ "bottom",
                               TRUE ~ NA)) %>% 
  ggplot(aes(factor(hidden_units), value_z, label = model_id)) +
  #geom_boxplot(width = .3, outlier.shape = NA) +
  geom_text(position = position_jitter(width = .1, height = 1.2)) +
  facet_wrap(vars(condition)) +
  theme_apa() +
  labs(x = "Hidden units", y = "Accuracy standardized (one point per model)", fill = "Hidden units") +
  #scale_fill_manual(values = c("firebrick4", "goldenrod3"))
  scale_color_brewer(palette = "Set2")

```

# Best ensembles across train, test, holdout

```{r}

TDF2024 %>% 
  filter(metric == "accuracy") %>%
  group_by(condition, model_id) %>% 
  summarise(acc = mean(value_z)) %>% 
  ggplot(aes(acc)) +
  geom_histogram(color = "black") +
  theme_apa() +
  facet_wrap(vars(condition))

```





## For abstract
### Mean differences


### Min max differences

```{r}

TDF2024 %>% 
  filter(hidden_units %in% c(10, 100)) %>% 
  group_by(metric, hidden_units) %>% 
  summarise(min_ = min(value_z),
            max_ = max(value_z)) %>% 
  mutate(accuracy_z_difference = min_ - max_)


```

### Descriptives, raw
```{r}
TDF2024 %>% 
  group_by(hidden_units, condition, metric) %>% 
  summarise(M = mean(value)) %>% 
  filter(condition == "train")


TDF2024 %>% 
  select(-accuracy_z, -mean_, -sd_) %>% 
  write_csv('~/Desktop/the_words_matter_subset_long.csv')

```
