---
title: "paper"
output: html_document
date: "2025-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 200)
require(tidyverse)
require(papaja)
require(RColorBrewer)
require(tools)
require(cowplot)
library(broom)
```

```{r data, include=FALSE}
words = read_csv('../../data/kidwords/kidwords.csv', col_names = c('word'))

# model summaries
model_summaries = read_csv('data/model_summaries.csv') %>% 
  pivot_longer(cols = accuracy_train:mse_holdout, names_to = "variable", values_to = "value") %>% 
  mutate(condition = case_when(str_detect(variable, "train") ~ "train",
                               str_detect(variable, "test") ~ "test",
                               str_detect(variable, "holdout") ~ "holdout"),
         metric = str_split(variable, "_", simplify = T)[,1]) %>% 
  filter(condition != "test") %>% 
  filter((hidden_units %in% c(20, 30, 40))) %>%
  dplyr::select(-variable)


model_summaries = model_summaries %>% 
  left_join(model_summaries %>% 
              group_by(condition, metric) %>% 
              summarise(M = mean(value),
                        SD = sd(value))) %>% 
  mutate(value_z = (value-M)/SD)


```

# Methods
## Frequencies
```{r frequencies}

probabilities = read_csv('data/word_frequencies_and_probabilities.csv') %>% 
  glimpse()

probabilities %>% 
  arrange(desc(frequency)) %>% 
  head()

```


# Results
## Differences in outcomes on training and generalization words
Means, standard deviations, and extreme values for raw and standardized MSE and raw accuracy for the train and generalization sets. Here we analyze data without respect to (collapsing across) representational capacity.

```{r descriptives1}
model_summaries %>% 
  filter(metric == "mse" | metric == "accuracy") %>% 
  group_by(condition, metric) %>% 
  filter(metric == "mse") %>% 
  summarise(mse_mean = mean(value),
            mse_sd = sd(value),
            mse_z_mean = mean(value_z),
            mse_z_sd = sd(value_z),
            mse_min = min(value),
            mse_z_min = min(value_z),
            mse_max = max(value),
            mse_z_max = max(value_z))

```

### Top and bottom performers

```{r}

bottom_100 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(condition) %>%
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  pull(model_id)

model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  filter(model_id %in% bottom_100)

# correlation
tmp  = model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank_train = seq_len(n())) %>% 
  select(tmp_id, rank_train)


tmp_2 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank_holdout = seq_len(n())) %>% 
  left_join(tmp) 

cor.test(tmp_2$rank_holdout, tmp_2$rank_train)

```

### Figure 2

```{r chunk3}
plot = model_summaries %>% 
  filter(metric == "mse") %>% 
  mutate(condition = factor(condition, levels = c("train", "holdout"))) %>% 
  ggplot(aes(factor(condition), value_z, fill = factor(condition))) +
  geom_violin() +
  geom_boxplot(width = .2) +
  theme_apa() +
  labs(x = "Set", y = "Mean squared error (Z)") +
  theme(text = element_text(family = "times"),
        legend.position = "none") +
  scale_x_discrete(labels = c("train" = "Train", "holdout" = "Generalization")) +
  scale_fill_manual(values = c("goldenrod3", "firebrick4"))


plot
ggsave("img/results_figure2.png", plot = plot, width = 5, height = 4, dpi = 600)

```


### Table 3

```{r}
model_summaries %>% 
  filter(metric == "mse") %>% 
  group_by(condition, hidden_units) %>% 
  summarise(mse_mean = mean(value),
          mse_sd = sd(value),
          mse_z_mean = mean(value_z),
          mse_z_sd = sd(value_z),
          mse_min = min(value),
          mse_z_min = min(value_z),
          mse_max = max(value),
          mse_z_max = max(value_z)) %>% 
  mutate(MSE = str_c(mse_mean, " (", mse_z_mean, ")"),
         SD = str_c(mse_sd, " (", mse_z_sd, ")"),
         MIN = str_c(mse_min, " (", mse_z_min, ")"),
         MAX = str_c(mse_max, " (", mse_z_max, ")")) %>% 
  select(hidden_units, MSE, SD, MIN, MAX)


```

### Figure 3

```{r}

plot = model_summaries %>% 
  filter(metric == "mse") %>% 
  mutate(condition = case_when(condition == "holdout" ~ "Generalization",
                               condition == "train" ~ "Train")) %>% 
  mutate(condition = factor(condition, levels = c("Train", "Generalization"))) %>% 
  ggplot(aes(factor(hidden_units), value_z, fill = interaction(factor(condition), factor(hidden_units)))) +
  geom_violin() +
  geom_boxplot(width = .5) +
  theme_apa() +
  facet_wrap(~condition) +
  theme(text = element_text(family = "times"),
        legend.position = "none") +
  labs(x = "Representational capacity (hidden units)",
       y = "Mean squared error (Z)") +
  scale_fill_manual(values = c("Train.20" = "lightgoldenrod", "Train.30" = "goldenrod3", "Train.40" = "orange", "Generalization.20" = "lightcoral", "Generalization.30" = "firebrick4", "Generalization.40" = "deeppink4"))

plot
ggsave("img/results_figure3.png", plot = plot, width = 5, height = 4, dpi = 600)


```

### Differences in Specific Ensembles at Different Levels of Representational Capacity
The correlation between 20 and 40 hidden units.

```{r}
tmp = model_summaries %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 20 & condition == "holdout") %>% 
  select(model_id, hu20 = value_z) %>% 
  left_join(model_summaries %>% 
  filter(metric == "mse") %>% 
  filter(hidden_units == 40 & condition == "holdout") %>% 
  select(model_id, hu40 = value_z))

cor.test(tmp$hu20, tmp$hu40)

```



#### Figure 4 combined plot

```{r}
text_size = 8

differences_in_rank = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(hidden_units) %>%
  #summarise(acc = mean(value_z)) %>% 
  arrange(desc(value_z)) %>% 
  mutate(rank = seq_len(n())) %>% 
  ungroup() %>%
  group_by(model_id) %>%
  arrange(-desc(hidden_units)) %>% 
  summarize(difference = diff(rank))


plot_a = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  left_join(differences_in_rank, by = "model_id") %>% 
  ggplot(aes(factor(hidden_units), value_z)) +
  geom_line(size = .2, aes(group = model_id, color = difference)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .2, outlier.size = .2) +
  theme_apa() +
  labs(title = "Differences in rank outcome\n 20, 40 hidden units (holdout only)",
       x = "Representational capacity (in hidden units)", 
       y = "Mean squared error (Z)", 
       fill = "Hidden units") +
  scale_fill_manual(values = c("lightcoral", "deeppink4")) +
  scale_color_gradientn(colors = heat.colors(10)) +
  theme(text = element_text(family = "times"),
        legend.position = "none",
        plot.title = element_text(size = text_size),
        plot.subtitle = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size))

plot_a


plot_b = differences_in_rank %>% 
  ggplot(aes(difference)) +
  geom_histogram(color = "black", fill = "turquoise") +
  labs(x = "Difference in rank outcome", y = "Count",
       title = "Differences in rank outcome for each model") +
  theme_apa() +
  theme(text = element_text(family = "times"),
        plot.title = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size))
  
plot_b  

# spotlighting top movers
bottom_10 = differences_in_rank %>% 
  arrange(difference) %>% 
  slice_head(n = 10) %>% 
  mutate(status = "bottom")

top_bottom_10 = differences_in_rank %>% 
  arrange(desc(difference)) %>% 
  slice_head(n = 10) %>% 
  mutate(status = "top") %>% 
  rbind(bottom_10)

plot_c = model_summaries %>% 
  filter(hidden_units %in% c(20, 40)) %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  left_join(top_bottom_10)  %>%
  mutate(model_id_ = case_when(is.na(status) ~ NA,
                               !is.na(status) ~ model_id)) %>% 
  ggplot(aes(factor(hidden_units), value_z)) +
  geom_violin(aes(fill = factor(hidden_units))) +
  geom_boxplot(width = .2, outlier.size = .2) +
  geom_line(size = .2, na.rm = TRUE, alpha = .9, aes(group = model_id_, color = status, linetype = status)) +
  theme_apa() +
  labs(title = "Top 20 movers in differences in outcome\nmodels at 20, 40 hidden units (holdout only)",
       x = "Representational capacity (in hidden units)", 
       y = "Mean squared error (Z)", 
       fill = "Hidden units") +
  scale_fill_manual(values = c("lightcoral", "deeppink4")) +
  #scale_color_gradientn(colors = heat.colors(10)) +
  theme(text = element_text(family = "times"),
        legend.position = "none",
        plot.title = element_text(size = text_size),
        plot.subtitle = element_text(size = text_size),
        axis.text = element_text(size = text_size),
        axis.title = element_text(size = text_size)) +
  scale_color_manual(values = c("top" = "grey44", "bottom" = "black")) +
  scale_linetype_manual(values = c("dashed", "dotted"))


plot_c

top<- plot_grid(plot_a, plot_c, ncol = 2)

# Arrange the bottom row plot centered
bottom <- plot_grid(NULL, plot_b, NULL, ncol = 3, rel_widths = c(.5, 1.5, .5))

# Combine the top and bottom rows
plot <- plot_grid(top, bottom, ncol = 1, rel_heights = c(1, 1))
plot


ggsave("img/results_figure4.png", plot = plot, width = 5, height = 4, dpi = 600)
```


