---
title: "Convergence"
output: html_document
date: "2025-05-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(arrow)
require(lme4)

CONTROL = lmerControl(optimizer = "bobyqa")



N = 500

set.seed(879)



twm = read_parquet("osf/data_for_lmem_full.parquet")
  
sampled_model_ids = sample(twm$model_id, N)
subsample = twm %>% 
  filter(holdout == TRUE) %>% 
  filter(model_id %in% sampled_model_ids)

item_ids = subsample %>% 
  distinct(word) %>% 
  rownames_to_column(var = "item_id")



noise_added_list = list()

SD = sd(twm$mse)
noise_added_list[["0"]] = subsample
for (i in seq(20)){
  
  tmp = subsample %>% 
    mutate(noise = rnorm(n(), mean = 0, sd = SD/1.6),
           mse = mse + noise,
           model_id = model_id + (i*10000)) %>% 
    select(-noise)
  
  print(cor(subsample$mse, tmp$mse))
  noise_added_list[[as.character(i)]] = tmp
  
}


noise_added = purrr::list_rbind(noise_added_list) %>% 
  glimpse() %>% 
  mutate(mse_z = (mse-mean(mse))/sd(mse)) 
```

## Try maximal model with dataset with noise added

```{r}

set.seed(765)

main_with_noise_added = noise_added %>% 
  lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL, verbose = 1)

main
summary(main)


```


## Test parameter estimates across different runs, manipulating seed

```{r}


#twm = read_parquet("osf/data_for_lmem_full.parquet") %>% 
#  filter(holdout == TRUE) %>% 
#  mutate(mse_z = (mse-mean(mse))/sd(mse))

  
models = list()
seedstart = 679
for (i in seq(20)){
  
  set.seed(seedstart + i)

  model = subsample %>% 
    mutate(mse_z = (mse-mean(mse))/sd(mse)) %>% 
    lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL, verbose = 2)
  
  print(paste("Model", i, "done"))
  
  models[[as.character(i)]] = model
  
}

# summaries
for (e in models){
  
  print(summary(e))
  
}

summary(model)
```

# N = 500; different random subsets
```{r}

N = 500
models = list()
seedstart = 679
for (i in seq(20)){
  
  set.seed(seedstart + i)
  sampled_model_ids = sample(twm$model_id, N)
  
  set.seed(782)
  model = twm %>%
    filter(model_id %in% sampled_model_ids) %>% 
    mutate(mse_z = (mse-mean(mse))/sd(mse)) %>% 
    lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL, verbose = 2)
  
  print(paste("Model", i, "done"))
  
  models[[as.character(i)]] = model
  
}

# summaries
for (e in models){
  
  print(summary(e))
  
}

summary(model)



```


# N = 1000; different random subsets
```{r}

N = 1000
models = list()
seedstart = 679
for (i in seq(20)){
  
  set.seed(seedstart + i)
  sampled_model_ids = sample(twm$model_id, N)
  
  set.seed(782)
  model = twm %>%
    filter(model_id %in% sampled_model_ids) %>% 
    mutate(mse_z = (mse-mean(mse))/sd(mse)) %>% 
    lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL, verbose = 2)
  
  print(paste("Model", i, "done"))
  
  models[[as.character(i)]] = model
  
}

# summaries
for (e in names(models)){
  
#  print(summary(e))
  outfile = str_c("osf/convergence/model_summary_n_1000_subsampled_model_id_", e, ".RData")
  model = models[e]
  save(model, file = outfile)

  
}

summary(model)



```


* I ran the above routine with different random seeds with the same subset of 500 randomly selected model IDs. All results were identical.
* I ran the above routine with different random subsets (random seed constant for lmer()) of 500 randomly selected model IDs. All results were very similar.
* I ran the above routine with different random subsets (random seed constant for lmer()) of 1000 randomly selected model IDs. All results were __________
