---
title: "LMEM"
output: html_document
date: "2025-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Call libraries
library(lattice)
library(Matrix)
library(lme4)
#library(MASS)
library(tidyverse)
library(lmerTest)
library(conflicted)



conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(lme4::lmer)


# control options of LMEM
# ftol_abs = 1e-8, xtol_abs = 1e-8, 

## fit models 1-2 with: rhobeg = 2e-3, rhoend = 2e-7
control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 5e6, rhobeg = .1, rhoend = 2e-9))

# data



twm = read_csv("data/subset_data_for_lmem_v2.csv") %>% 
  filter(holdout == TRUE) %>% 
  filter(hidden %in% c(20, 30, 40)) %>%
  glimpse() %>% 
  mutate(mse_z = (mse-mean(mse))/sd(mse)) 

stopifnot(all(twm$holdout)) # should be holdout items only

```

# Linear Mixed Effects Models

## Statistical Model 1: Unconditional Model A
Here we have our unconditional without model_id (nested within unconditional)
```{r StatisticalModel1}
uncon_no_model_id = lmer(mse_z ~ 1 + (1 | word), data = twm, control = control)
summary(uncon_no_model_id)
```


## Statistical Model 2: Unconditional Model B
This is the unconditional model (nested within main_no_hidden_slopes)
```{r StatisticalModel2}

uncon = lmer(mse_z ~ 1 + (1 | model_id) + (1 | word), data = twm, control = control)

uncon = lmer(mse_z ~ 1 + (1 | model_id) + (1 | word), data = twm)
uncon_allfits = allFit(uncon)

summary(uncon)

```


## Statistical Model 3: Main Model No Slopes
This is the main model without any slopes (nested within model_revised)

```{r StatisticalModel3}
# 
CONTROL = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 5e6, rhobeg = .1, rhoend = 2e-9))
main_no_hidden_slopes = twm %>% 
  mutate(hidden = hidden/10) %>% 
  lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL)

summary(main_no_hidden_slopes)


```


## Statistical Model 4: Main Model No Hidden Slope on Ensemble
This is the main model without hidden slope on model_id (nested within main)

```{r StatisticalModel4}
main_revised = twm %>% 
  mutate(hidden = hidden/10) %>% 
  lmer(mse_z ~ 1 + hidden + (1 |model_id) + (1 + hidden | word), data = ., control = CONTROL)

summary(main_revised)
```


## Statistical Model 5: Full Model
The main model (full model)

```{r StatisticalModel5}


main = twm %>% 
  mutate(hidden = hidden/1000) %>% 
  lmer(mse_z ~ 1 + hidden + (1 + hidden | model_id) + (1 + hidden | word), data = ., control = CONTROL)


save(main, file = "main_model.RData")
summary(main)

```


```{r}

iterations = 20
models = list()

for (iter in seq(iterations)){
  
  extra = twm %>% 
  #mutate(hidden = hidden) %>% 
  lmer(mse_z ~ 1 + hidden + (1 + hidden | model_id/word), data = ., control = CONTROL)
  
  id = paste("model_", iter, sep = "")
  models[[id]] = extra
  
}

for (m in models){
  
  print(summary(m))
  
  
}
summary(extra)




```

## Crossed
```{r}

iterations = 5
models_main = list()

for (iter in seq(iterations)){
  
  extra_ = twm %>% 
    mutate(hidden = hidden/10000) %>% 
    lmer(mse_z ~ 1 + hidden + (1 + hidden || model_id) + (1 + hidden || word), data = ., control = CONTROL)
  
  print(paste("Iteration", iter, "done"))
  id = paste("model_", iter, sep = "")
  models_main[[id]] = extra_
  
}

for (m in models_main){
  
  print(summary(m))
  
  
}

summary(extra_)



```

### One last try but with correlations of random effects


```{r}
extra_2 = twm %>% 
    mutate(hidden = hidden/10000) %>% 
    lmer(mse_z ~ 1 + hidden + (1 + hidden | model_id) + (1 + hidden | word), data = ., control = CONTROL)


```


```{r significanceTests}
anova(uncon_no_model_id, uncon) # is model_id necessary? (Does performance vary based on ensemble?)
anova(uncon, main_no_hidden_slopes) # is fixed effect of hidden necessary? (Does representational capacity affect MSE?)
anova(main_no_hidden_slopes, main_revised) # is hidden slope on word necessary? (Does the word variability change based on representational capacity?)
anova(main, main_revised) # is hidden slope on model_id necessary? (Does the ensemble variability change based on representational capacity?)

```

