---
title: "LMEM"
output: html_document
date: "2025-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Call libraries
require(arrow)
library(lattice)
library(Matrix)
library(lme4)
#library(MASS)
library(tidyverse)
library(lmerTest)
library(conflicted)



conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(lme4::lmer)


# control options of LMEM
CONTROL = lmerControl(optimizer = "bobyqa")

# data
twm = read_parquet("osf/data_for_lmem_full.parquet") %>% 
  filter(holdout == TRUE) %>% 
  glimpse() %>% 
  mutate(mse_z = (mse-mean(mse))/sd(mse),
        hidden_standardized = if_else(hidden == 40, 1, if_else(hidden == 20, -1, 0))) 

davg <- twm %>%
    group_by(model_id) %>%
    summarize(mse_mean = mean(mse)) %>%
    mutate(mse_rank = rank(mse_mean)) %>%
    mutate(strata = cut(mse_rank, seq(0, 10000, by = 1000))) %>%
    group_by(strata)

dsubset <- davg %>%
    filter((mse_rank %% 100) == 0) %>%
    arrange(mse_rank)

subset <- twm %>%
    filter(
        model_id %in% dsubset$model_id
    ) %>%
    mutate(
        mse = (mse - mean(mse)) / sd(mse))

stopifnot(all(twm$holdout)) # should be holdout items only

```

# Linear Mixed Effects Models

## Statistical Model 1: Unconditional Model A
Here we have our unconditional without model_id (nested within unconditional)
```{r StatisticalModel1}
set.seed(765)

unconditional_no_model_id = subset %>% 
  lmer(mse_z ~ 1 + (1 | word), data = ., control = CONTROL)

save(unconditional_no_model_id, file = "osf/unconditional_no_model_id.RData")

summary(unconditional_no_model_id)
```


## Statistical Model 2: Unconditional Model B
This is the unconditional model (nested within main_no_hidden_slopes)
```{r StatisticalModel2}

set.seed(765)

unconditional = subset %>% 
 lmer(mse_z ~ 1 + (1 | model_id) + (1 | word), data = ., control = CONTROL)

save(unconditional, file = "osf/unconditional.RData")

summary(unconditional)
```


## Statistical Model 3: Main Model No Slopes
This is the main model without any slopes (nested within model_revised)

```{r StatisticalModel3}
set.seed(765)

main = subset %>% 
  lmer(mse_z ~ 1 + hidden_standardized + (1 | model_id) + (1 | word), data = ., control = CONTROL)

main
summary(main)

save(main, file = "osf/main.RData")


```


### Model comparisons
An unconditional model with random intercepts for word and ensemble had a better fit than a model without an ensemble random effect.
```{r significanceTest1}
model_comparison_1 = anova(unconditional_no_model_id, unconditional) # is model_id necessary? (Does performance vary based on ensemble?)

save(model_comparison_1, file = "osf/model_comparison_1.RData")


# is fixed effect of hidden necessary? (Does representational capacity affect MSE?)
model_comparison_2 = anova(unconditional, main) 
save(model_comparison_2, file = "osf/model_comparison_2.RData")

```
