---
title: "LMEM"
output: html_document
date: "2025-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Call libraries
library(lattice)
library(Matrix)
library(lme4)
#library(MASS)
library(tidyverse)
library(lmerTest)
library(conflicted)



conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(lme4::lmer)


# control options of LMEM
CONTROL = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 5e6, rhobeg = .1, rhoend = 2e-9))

# data
twm = read_csv("data/subset_data_for_lmem_v2.csv") %>% 
  filter(holdout == TRUE) %>% 
  filter(hidden %in% c(20, 30, 40)) %>%
  glimpse() %>% 
  mutate(mse_z = (mse-mean(mse))/sd(mse)) 

stopifnot(all(twm$holdout)) # should be holdout items only

```

# Linear Mixed Effects Models

## Statistical Model 1: Unconditional Model A
Here we have our unconditional without model_id (nested within unconditional)
```{r StatisticalModel1}
set.seed(765)

uncon_no_model_id = twm %>% 
  lmer(mse_z ~ 1 + (1 | word), data = ., control = CONTROL)

save(uncon_no_model_id, file = "data/uncon_no_model_id.RData")

summary(uncon_no_model_id)
```


## Statistical Model 2: Unconditional Model B
This is the unconditional model (nested within main_no_hidden_slopes)
```{r StatisticalModel2}

set.seed(765)

uncon = twm %>% 
 lmer(mse_z ~ 1 + (1 | model_id) + (1 | word), data = ., control = CONTROL)

save(uncon, file = "data/uncon.RData")


summary(uncon)

```


## Statistical Model 3: Main Model No Slopes
This is the main model without any slopes (nested within model_revised)

```{r StatisticalModel3}

set.seed(765)

main_no_hidden_slopes = twm %>% 
  lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 | word), data = ., control = CONTROL)

main_no_hidden_slopes
summary(main_no_hidden_slopes)

save(main_no_hidden_slopes, file = "data/main_no_hidden_slopes.RData")


```


## Statistical Model 4: Main Model No Hidden Slope on Ensemble
This is the main model without hidden slope on model_id (nested within main)

```{r StatisticalModel4}
set.seed(765)

main_revised = twm %>% 
  lmer(mse_z ~ 1 + hidden + (1 | model_id) + (1 + hidden | word), data = ., control = CONTROL)

main_revised
save(main_revised, file = "data/main_revised.RData")

summary(main_revised)
```


## Statistical Model 5: Full Model
The main model (full model)

```{r StatisticalModel5}
set.seed(765)

main = twm %>% 
  lmer(mse_z ~ 1 + hidden + (1 + hidden | model_id) + (1 + hidden| word), data = ., control = CONTROL)


save(main, file = "data/main_model.RData")
summary(main)

```

### Model comparisons
An unconditional model with random intercepts for word and ensemble had a better fit than a model without an ensemble random effect.
```{r significanceTest1}
model_comparison_1 = anova(uncon_no_model_id, uncon) # is model_id necessary? (Does performance vary based on ensemble?)

save(model_comparison_1, file = "data/model_comparison_1.RData")



# additional model comparisons for reference:
model_comparison_2 = anova(uncon, main_no_hidden_slopes) # is fixed effect of hidden necessary? (Does representational capacity affect MSE?)
save(model_comparison_2, file = "data/model_comparison_2.RData")


model_comparison_3 = anova(main_no_hidden_slopes, main_revised) # is hidden slope on word necessary? (Does the word variability change based on representational capacity?)
save(model_comparison_3, file = "data/model_comparison_3.RData")
```

Comparison of model with a slope for hidden units in model ID ("ensemble") against one without such a slope. The more complex statistical model (with the hidden unit slope for model ID) fit better.
```{r significanceTest2}


model_comparison_final = anova(main, main_revised) # is hidden slope on model_id necessary? (Does the ensemble variability change based on representational capacity?)
save(model_comparison_final, file = "data/model_comparison_final.RData")

```



