---
title: "paper"
output: html_document
date: "2025-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 200)
require(tidyverse)
require(papaja)
require(RColorBrewer)
require(tools)
require(cowplot)
library(broom)
```

```{r data, include=FALSE}
source("load.R")


```

# Methods
## Frequencies
```{r frequencies}

probabilities = read_csv('data/word_frequencies_and_probabilities.csv') %>% 
  glimpse()

probabilities %>% 
  arrange(desc(frequency)) %>% 
  head()

```


# Results
## Differences in outcomes on training and generalization words
Means, standard deviations, and extreme values for raw and standardized MSE and raw accuracy for the train and generalization sets. Here we analyze data without respect to (collapsing across) representational capacity.

```{r descriptives1}
model_summaries %>% 
  filter(metric == "mse" | metric == "accuracy") %>% 
  group_by(condition, metric) %>% 
  filter(metric == "mse") %>% 
  summarise(mse_mean = mean(value),
            mse_sd = sd(value),
            mse_z_mean = mean(value_z),
            mse_z_sd = sd(value_z),
            mse_min = min(value),
            mse_z_min = min(value_z),
            mse_max = max(value),
            mse_z_max = max(value_z))

```

### Identify top and bottom 100 performers
The number of rows resulting indicates the number of bottom 100 performers in train that were also in the bottom 100 performers in holdout.
```{r}

bottom_100 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(condition) %>%
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  pull(model_id)

model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  filter(model_id %in% bottom_100)

```



### Table 3

```{r}
model_summaries %>% 
  filter(metric == "mse") %>% 
  group_by(condition, hidden_units) %>% 
  summarise(mse_mean = round(mean(value), digits = 3),
          mse_sd = sd(value),
          mse_z_mean = mean(value_z),
          mse_z_sd = round(sd(value_z), digits = 3),
          mse_min = min(value),
          mse_z_min = round(min(value_z), digits = 2),
          mse_max = max(value),
          mse_z_max = round(max(value_z), digits = 3)) %>% 
  mutate(MSE = str_c(mse_mean, " (", mse_z_mean, ")"),
         SD = str_c(mse_sd, " (", mse_z_sd, ")"),
         MIN = str_c(mse_min, " (", mse_z_min, ")"),
         MAX = str_c(mse_max, " (", mse_z_max, ")")) %>% 
  select(hidden_units, MSE, SD, MIN, MAX)


```
