---
title: "paper"
output: html_document
date: "2025-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 200)
require(tidyverse)
require(papaja)
require(RColorBrewer)
require(tools)
require(cowplot)
library(broom)
```

```{r data, include=FALSE}
source("load.R")


```

# Intro
No tables calculated in this script. See Table 1 CSV for that data, which is calculated offline.

# Methods
No tables calculated in this script. See Table 2 CSV for that data, which is calculated offline.

# Results
## Table 3
Means, standard deviations, and extreme values for raw and standardized MSE and raw accuracy for the train and generalization sets. Here we analyze data without respect to (collapsing across) representational capacity. These data populate Table 3.

```{r descriptives1}
model_summaries %>% 
  filter(metric == "mse" | metric == "accuracy") %>% 
  group_by(condition, metric) %>% 
  filter(metric == "mse") %>% 
  summarise(mse_mean = mean(value),
            mse_sd = sd(value),
            mse_z_mean = mean(value_z),
            mse_z_sd = sd(value_z),
            mse_min = min(value),
            mse_z_min = min(value_z),
            mse_max = max(value),
            mse_z_max = max(value_z))

```

### Identify top and bottom 100 performers
The number of rows resulting indicates the number of bottom 100 performers in train that were also in the bottom 100 performers in holdout.
```{r}

bottom_100 = model_summaries %>% 
  filter(metric == "mse" & condition == "holdout") %>% 
  group_by(condition) %>%
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  pull(model_id)

model_summaries %>% 
  filter(metric == "mse" & condition == "train") %>% 
  mutate(tmp_id = str_c(model_id, "_", hidden_units)) %>% 
  arrange(desc(value)) %>% 
  mutate(rank = seq_len(n())) %>% 
  filter(rank < 100) %>% 
  filter(model_id %in% bottom_100)

```

## Other tables
Tables 4 and 5 are calculated within the primary manuscript script.